Report for Qualifying Exam

### 2024-07-31 src/sections/introduction.tex
added SVG, fixed table

\section{Introduction} \label{sec:introduction}
In the digital age, online information seekers frequently encounter \emph{search results} laden with \emph{misinformation}, \emph{biased narratives}, and \emph{conspiracy theories}.
This exposure can potentially lead users to accept and propagate false information.
However, this phenomenon is not universal across all search queries.
Certain searches, known as \textit{data voids} are particularly susceptible to these issues.
\\
\textit{Data voids} occur when searching for terms that yield limited, non-existent, or highly problematic relevant information.
Unlike common searches that produce abundant data, queries falling into data voids may return no results or present irrelevant and often inaccurate information.
Data voids can be mainly categorized into two main types:
\begin{itemize}
    \item newly coined terms with no established information base, and
    \item existing terms whose meanings have evolved over time.
\end{itemize}
This report focuses on the latter, examining data voids where context drift occurs.
The following papers has been selected to explore this phenomenon.
\begin{itemize}
    \item Statistically Significant Detection of Linguistic Change./ ACM WWW 2014 ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic}
    \item Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change./ ACL 2016 ~\cite{hamilton-etal-2016-diachronic}
    \item The Pandemic in Words: Tracking Fast Semantic Changes via a Large-Scale Word Association Task. \cite{10.1162/opmi_a_00081}
\end{itemize}

>>>>

\section{Introduction} \label{sec:introduction}
In the digital age, online information seekers frequently encounter \emph{search results} laden with \emph{misinformation}, \emph{biased narratives}, and \emph{conspiracy theories}.
This exposure can potentially lead users to accept and propagate false information.
However, this phenomenon is not universal across all search queries.
Certain searches, known as \textit{data voids} are particularly susceptible to these issues.
\\
\textit{Data voids} occur when searching for terms that yield limited, non-existent, or highly problematic relevant information.
Unlike common searches that produce abundant data, queries falling into data voids may return no results or present irrelevant and often inaccurate information.
Data voids can be mainly categorized into two main types:
\begin{itemize}
    \item newly coined terms with no established information base, and
    \item existing terms whose meanings have evolved over time.
\end{itemize}
This report focuses on the latter, examining data voids where context drift occurs.
The following papers has been selected to explore this phenomenon.
\begin{itemize}
    \item Statistically Significant Detection of Linguistic Change.
    ACM WWW 2014 ~\cite{kulkarni2014statisticallysignificantdetectionlinguistic}
    \item Diachronic Word Embeddings Reveal Statistical Laws of Semantic Change.
    ACL 2016 ~\cite{hamilton-etal-2016-diachronic}
\end{itemize}

### 2024-07-31 src/sections/relatedwork.tex
added SVG, fixed table

\multicolumn{1}{c}{\textbf{\begin{tabular}[c]{@{}c@{}}Linguistic \\ Change\end{tabular}}} & \multicolumn{1}{c}{\textbf{Approach}}                                              & \multicolumn{1}{c}{\textbf{TCD}} & \multicolumn{3}{c}{\textbf{Example}}                                                                                                                                                                 \\ \cmidrule(l){4-6}
\multicolumn{1}{c}{\textbf{}}                                                             & \multicolumn{1}{c}{\textbf{}}                                              & \multicolumn{1}{c}{\textbf{}}    & \textbf{Word}         & \textbf{Former Usage}                                                                & \textbf{New Usage}                                                                    \\ \midrule
\multicolumn{1}{l}{\textbf{Semantic}}                                                     & \multicolumn{1}{l}{\textbf{FREQUENCY}}                                             & \multicolumn{1}{l}{}            & \multicolumn{1}{l}{} & \multicolumn{1}{l}{}                                                                & \multicolumn{1}{l}{}                                                                 \\ \midrule
\textbf{}                                                                                 & Log Ratio                                                                  & 1960 \& 1990                     & disk                  & -                                                                                    & -                                                                                     \\ \cmidrule(l){2-6}
\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Word \\ Frequency\end{tabular}                  & 1900--2005                      & bitch                 & female dog                                                                           & slang                                                                                 \\ \cmidrule(l){2-6}
\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Law of \\ Conformity\end{tabular}               & 1800--2000                      & -                     & -                                                                                    & -                                                                                     \\ \cmidrule(l){2-6}
                                                                                          & \textbf{DISTRIBUTIONAL}                                                             &                                  &                       &                                                                                      &                                                                              \\ \cmidrule(l){2-6}

@@

\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Continuous \\ Word \\ Embeddings\end{tabular}   & 1900--2009                      & cell                  & closet, dungeon                                                                      & phone, cordless                                                                       \\ \cmidrule(l){2-6}

@@

\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Cultural and \\ linguistic Drift\end{tabular}   & 1800--2000                      & virus                 & \begin{tabular}[c]{@{}l@{}}infected with\\  the virus\end{tabular}                   & \begin{tabular}[c]{@{}l@{}}spreading\\  computer virus\end{tabular}                   \\ \cmidrule(l){2-6}
\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Diachronic \\ Word\\ Embeddings\end{tabular}    & 1800--1999                      & awful                 & full of awe                                                                          & terrible or appalling                                                                 \\ \cmidrule(l){2-6}
\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Contextual \\ Word \\ Embeddings\end{tabular}                                                                      & 1910 – 2009                      & tenure                & \begin{tabular}[c]{@{}l@{}}short term leases,\\ insecurity of tenure\end{tabular}    & \begin{tabular}[c]{@{}l@{}}tenure of office, \\ employment \\ and tenure\end{tabular} \\ \cmidrule(l){2-6}
\textbf{}                                                                                 & \begin{tabular}[c]{@{}l@{}}Semantic \\ Similarity \\ Analysis\end{tabular} & 2014--2022                      & immunity              & \begin{tabular}[c]{@{}l@{}}politics (‘legislator’,\\ ‘representative’).\end{tabular} & \begin{tabular}[c]{@{}l@{}}health-related \\ (‘prevention’, ‘AIDS’)\end{tabular}      \\ \cmidrule(l){2-6}
                                                                                          & \textbf{PART OF SPEECH}                                                                        &                                  &                       &                                                                                      &                                                                   \\ \midrule

@@

\bottomrule
\end{tabular}
%\bottomrule
\caption{Various types of linguistic change and method used to detect them.}
\label{tab:related-work}
\end{table}

% ------------------------------------------------------------------ SEMANTIC ------------------------------------------------------------------
\begin{itemize}
    \item \textbf{Investigating Semantic Shifts Through Word Contexts.}
    Gulordava and Baroni (2011) detected semantic change by focusing on words used in the 1960s and 1990s.
        They compared the similarity of the surrounding words (words that co-occur with the target word) in these two time periods.
    \begin{itemize}
        \item To assess similarity, the researchers calculated the \emph{Local Mutual Information (LMI)} score between the central word and its surrounding words.
        A low LMI score between the target word and its surrounding words across the two time periods indicated a potential semantic shift.
        \item The study demonstrated that their distributional similarity models were effective in capturing \emph{cultural shifts} in word meaning.
        For example, they found that the word `sleep' acquired more negative connotations related to sleep disorders when comparing its contexts in the 1960s to those in the 1990s.
    \end{itemize}

@@

    \item \para{Semantic Similarity:} The impact of significant events like the COVID-19 pandemic on language and semantic change has also been a subject of study.
        Laurino et al.\ (2023) explores tracking fast semantic changes through a large-scale word association task, aiming to understand how the collective mental lexicon evolves in response to such global events.
        Their research highlights the dynamic nature of language and how it incorporates new senses.
        For instance, words like `quarantine', `mask', and `social distancing' took on new and prominent meanings in everyday conversation.
\end{itemize}

@@

% ------------------------------------------------------------------ THREE PAPERS ------------------------------------------------------------------
Several studies from our literature review provide valuable insights for addressing data voids in our analysis.
\begin{itemize}
    \item The first being the work by Kulkarni et al.\ which introduces us to time series construction.
Their distributional methods focuses on finding subtle semantic shits to determine the context where a word in used.
This concept aligns perfectly with our goal of uncovering shifts in word usage when encountering data voids.
    \item Second paper by Hamilton et al.\ introduces us to word emebeddings alignment.
Since we want to compare word vectors from different time periods, vectors should be aligned in same coordinate axes.
After aligning the embeddings for individual time periods, we can use the aligned word vectors to compute the semantic displacement that a word has undergone during a certain time-period.
    \item Third one by Laurino et al.\ investigates the impact of the COVID-19 pandemic on word meaning.
One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.
This suggests that these words underwent a more significant and rapid semantic shift compared to control words not associated with the pandemic.
They also employ semantic similarity analysis to quantify the shifts in meaning for pandemic-related words and provides evidence that the COVID-19 pandemic acted as a catalyst for rapid semantic change.
\end{itemize}

>>>>

\multicolumn{1}{c}{\textbf{Linguistic}}              & \multicolumn{1}{c}{\textbf{Approach}}                                            & \multicolumn{1}{c}{\textbf{TCD}}      & \multicolumn{3}{c}{\textbf{Example}}                                                                                                                                                                              \\ \cmidrule(l){4-6}
\multicolumn{1}{c}{\textbf{Change}}                  & \multicolumn{1}{c}{\textbf{}}                                                    & \multicolumn{1}{c}{\textbf{}}         & \textbf{Word}             & \textbf{Former Usage}                                                                         & \textbf{New Usage}                                                                    \\ \midrule
\rowcolor{lightgray}                                 & \multicolumn{5}{c}{\textbf{FREQUENCY \cite{gulordava-baroni-2011-distributional, kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-diachronic}}}                                                                                                                                                                   \\
\textbf{\textbf{Semantic}}                           & Log Ratio                                                                        & 1960 \& 1990                          & disk                      & -                                                                                             & -                                                                                     \\ \cmidrule(l){2-6}
\textbf{}                                            & Word Frequency                                                                   & 1900--2005                            & bitch                     & female dog                                                                                    & slang                                                                                 \\ \cmidrule(l){2-6}
\rowcolor{lightgray}                                 & \multicolumn{5}{c}{\textbf{DISTRIBUTIONAL \cite{gulordava-baroni-2011-distributional, kim-etal-2014-temporal, kulkarni2014statisticallysignificantdetectionlinguistic, 10.1007/978-3-319-50496-4_18, hamilton-etal-2016-diachronic, hamilton-etal-2016-cultural, 10.1162/opmi_a_00081, giulianelli-etal-2020-analysing}}}                                                                                                                                         \\ \cmidrule(l){2-6}

@@

\textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Continuous Word \\ Embeddings\end{tabular}            & 1900--2009                            & cell                      & closet, dungeon                                                                               & phone, cordless                                                                       \\ \cmidrule(l){2-6}

@@

\textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Cultural and \\ Linguistic Drift\end{tabular}         & 1800--2000                            & virus                     & \begin{tabular}[c]{@{}l@{}}infected with virus\end{tabular}                                   & \begin{tabular}[c]{@{}l@{}}spreading\\  computer virus\end{tabular}                   \\ \cmidrule(l){2-6}
\textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Diachronic Word\\ Embeddings\end{tabular}             & 1800--1999                            & awful                     & full of awe                                                                                   & terrible or appalling                                                                 \\ \cmidrule(l){2-6}
\textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Contextual Word \\ Embeddings\end{tabular}            & 1910 – 2009                           & tenure                    & \begin{tabular}[c]{@{}l@{}}short term leases,\\ insecurity of tenure\end{tabular}             & \begin{tabular}[c]{@{}l@{}}tenure of office, \\ employment \end{tabular}              \\ \cmidrule(l){2-6}
\textbf{}                                            & \begin{tabular}[c]{@{}l@{}}Semantic Similarity \\ Analysis\end{tabular}          & 2014--2022                            & immunity                  & \begin{tabular}[c]{@{}l@{}}politics (‘legislator’,\\ ‘representative’).\end{tabular}          & \begin{tabular}[c]{@{}l@{}}health-related \\ (‘prevention’, ‘AIDS’)\end{tabular}      \\ \cmidrule(l){2-6}
\rowcolor{lightgray}                                 & \multicolumn{5}{c}{\textbf{PART OF SPEECH \cite{kulkarni2014statisticallysignificantdetectionlinguistic, hamilton-etal-2016-cultural}}}                                                                                                                                                                                                      \\

@@

\bottomrule
\end{tabular}

\caption{\textbf{Comparative Overview of Linguistic Change Detection Approaches.}
This table provides a comprehensive summary of various approaches and contributions to the study of linguistic changes,
    categorizing the type of changes (semantic or syntactic), and the methodologies used, including frequency and distributional approaches.
    It highlights each study’s methods, such as word embeddings, statistical analysis, and clustering, along with the time periods of change detection (TCD) and examples of words.}
\label{tab:related-work}
\end{table}
\vspace{0mm}
% ------------------------------------------------------------------ SEMANTIC ------------------------------------------------------------------
\begin{itemize}
    \item \textbf{Investigating Semantic Shifts Through Word Contexts.}
    Gulordava and Baroni (2011) detected semantic change by focusing on words used in the 1960s and 1990s.
        They compared the similarity of the surrounding words (words that co-occur with the target word) in these two time periods.
    \begin{itemize}
        \item To assess similarity, the researchers calculated the \emph{Local Mutual Information (LMI)} score between the central word and its surrounding words.
        A low LMI score between the target word and its surrounding words across the two time periods indicated a potential semantic shift.
        \item The study demonstrated that their distributional similarity models were effective in capturing \emph{cultural shifts} in word meaning.
        For example, they found that the word `sleep' acquired more negative connotations related to sleep disorders when comparing its contexts in the 1960s to those in the 1990s.
    \end{itemize}

@@

    \item \para{Semantic Similarity:} The impact of significant events like the COVID-19 pandemic on language and semantic change has also been a subject of study.
    \begin{itemize}
        \item Laurino et al.\ (2023) explores tracking fast semantic changes through a large-scale word association task, aiming to understand how the collective mental lexicon evolves in response to such global events.
            Their research highlights the dynamic nature of language and how it incorporates new senses.
        \item One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.
            This suggests that these words underwent a more significant and rapid semantic shift compared to control words not associated with the pandemic.
            They also employ semantic similarity analysis to quantify the shifts in meaning for pandemic-related words and provides evidence that the COVID-19 pandemic acted as a catalyst for rapid semantic change.
            For instance, words like `quarantine', `mask', and `social distancing' took on new and prominent meanings in everyday conversation.
    \end{itemize}
\end{itemize}

@@

% ------------------------------------------------------------------ THREE PAPERS ------------------------------------------------------------------
Several studies from our literature review provide valuable insights for addressing data voids in our analysis.
\begin{itemize}
    \item The first being the work by Kulkarni et al.\ which introduces us to time series construction.
Their distributional methods focuses on finding subtle semantic shits to determine the context where a word in used.
This concept aligns perfectly with our goal of uncovering shifts in word usage when encountering data voids.
    \item Second paper by Hamilton et al.\ introduces us to word emebeddings alignment.
Since we want to compare word vectors from different time periods, vectors should be aligned in same coordinate axes.
After aligning the embeddings for individual time periods, we can use the aligned word vectors to compute the semantic displacement that a word has undergone during a certain time-period.
%    \item Third one by Laurino et al.\ investigates the impact of the COVID-19 pandemic on word meaning.
%One key finding from their work is that words directly related to the pandemic exhibited a greater difference in semantic similarity between pre-pandemic and pandemic time periods.
%This suggests that these words underwent a more significant and rapid semantic shift compared to control words not associated with the pandemic.
%They also employ semantic similarity analysis to quantify the shifts in meaning for pandemic-related words and provides evidence that the COVID-19 pandemic acted as a catalyst for rapid semantic change.
\end{itemize}
